---
title: "Activity Analysis"
output: md_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r Load Package}
library(dplyr)
library(chron)
library(ggplot2)
```

```{r Read data}
activity_data<-read.csv("activity.csv")
```

```{r Preprocess data}
activity_data$date<-as.Date(activity_data$date,"%Y-%m-%d")
```

```{r Calculate Mean}
#Remove NA values
activity_data1<-activity_data[complete.cases(activity_data),]
a<-activity_data1%>%group_by(date)%>%summarize(steps=sum(steps))
hist(a$steps,xlab = "Steps",ylab = "Frequency",main=" Histogram of Total number of steps taken each day")
a1<-mean(a$steps)
a2<-median(a$steps)
```
##Mean  of the total number of steps taken per day
```{r}
a1
```
##Median of the total number of steps taken per day
```{r}
a2
```
##Average daily activity pattern
```{r}
a3<-activity_data1%>%group_by(interval)%>%summarize(steps=sum(steps))
plot(a3$interval,a3$steps,type="l",xlab="Interval",ylab = "Steps",main = "Steps vs Time Interval")
```

# Interval containing the maximum number of steps
```{r}
a3[a3$steps==max(a3$steps),"interval"]
```
##Total number of missing values in the dataset
```{r}
nrow(activity_data[is.na(activity_data$steps),])
```
##Fill missing values with mean of that day
```{r}
a4<-activity_data1%>%group_by(interval)%>%summarize(steps=mean(steps))
merged_data<-merge(activity_data,a4,by.x = "interval",by.y = "interval")
merged_data$steps<-ifelse(is.na(merged_data$steps.x),merged_data$steps.y,merged_data$steps.x)
```
##New dataset that is equal to the original dataset but with the missing data filled in
```{r}
new_dataset<-merged_data[,c("date","interval","steps")]
head(new_dataset)
```
##histogram of the total number of steps taken each day 
```{r}
new_hist<-new_dataset%>%group_by(date)%>%summarize(steps=sum(steps))
hist(new_hist$steps,xlab="Steps",ylab="Frequency",main = "Histogram  Total number of steps taken each day")
```
##Mean of total number of steps taken per day
```{r}
mean(new_hist$steps)
```
##Median of total number of steps taken per day
```{r}
median(new_hist$steps)
```
## Impact of imputing missing data on the estimates of the total daily number of steps
```{r}
#For Mean
mean(new_hist$steps)-a1
#For Median
median(new_hist$steps)-a2
```
##Differences in activity patterns between weekdays and weekends
```{r}
new_dataset$day<-ifelse(is.weekend(new_dataset$date),"weekend","weekday")
new_plot<-new_dataset%>%group_by(interval,day)%>%summarize(steps=mean(steps))
ggplot(new_plot,aes(x=interval,y=steps))+geom_line()+facet_wrap(~day)+labs(x="Interval",y="Steps")
```
